\large{\bf Looking forwards and backwards through locally regulated populations }

### 2022-08-01 forwards-and-backwards.tex
fixed up proof of boundedness

\begin{definition}[Model parameterization and assumptions]
\label{def:model_setup}
Suppose $\gamma(x, m)$, $r(x, m)$, and $F(x, m)$
are uniformly continuous functions
from $\IR^d \times \IR_{\ge 0}$ to $\IR$.
Furthermore, suppose $r$ takes values in $[0, 1]$
and has uniformly bounded second derivatives,
that $\gamma(x,m) \ge 0$,
and that $m^2 \gamma(x, m)$ is uniformly bounded.
Fix positive values $\theta$, $\epsilon_r$, $\epsilon_\gamma$, and $\epsilon_F$,
with $\epsilon_r < \epsilon_\gamma$.
Suppose $F(x,m)$ is bounded above
with uniformly bounded first and second derivatives in $m$,
and define
\begin{align} \label{eqn:mu_defn}
    \mu(x, \eta)
    =
    \max\left(0, 
        r(x, \smooth{r} \eta(x))
        \gamma(\smooth{\gamma} \eta(x))
        - \frac{1}{\theta}
        F(x, \smooth{F} \eta(x)) 
    \right) .
\end{align}
Although the models will make sense without this requirement,
for the purposes of convergence we will also require that $\theta > 1/(\epsilon_\gamma - \epsilon_r)$.
\end{definition}

@@

\begin{proof}[Proof of Theorem~\ref{thm:nonlocal_convergence}.]
Here, we need to prove that for any sequence $N_k \to \infty$,
there is a subsequence $N_{k(j)}$ along which $(\eta^{N_{k(j)}}_t)_{t\ge 0}$ converges in distribution
(i.e., tightness for the family of distributions on c\`adl\`ag measure-valued paths),
and that any limit is a solution to the martingale problem stated in the Theorem.
This would convergence along $N_k$ to a unique limit
if the solution to the martingale problem was unique;
however, we only know this to be the case in certain situations.

@@

To prove tightness of the population process $\eta_t$,
we first show that the expected total mass at some fixed time $T$
is controlled uniformly across $N$, a compact containment condition.
To establish this control,
we first prove a technical lemma which controls spatial density of our population,
and then we use the Burkholder-Davis-Gundy inequality \comment{TODO CITE}
to give us control over the running supremum of the total population mass.
Once we have control over the total mass of the population, 
we will also have control over local densities.
This then allows us to prove that the particle system is tight
by an application of the Aldous-Rebolledo criterion \comment{TODO CITE}.

\comment{summarize and move to the appendix}

\subsubsection{Compact Containment Condition}
Before proving compact containment condition,
we first prove a technical lemma that involves a bound on the Gaussian kernels.
\begin{lemma}
    \label{lem: gamma - r radius comparison}
    Under the assumptions of Definition~\ref{def:model_setup},
    if $\epsilon_{r}^2 + \frac{1}{\theta} < \epsilon_{\gamma}^2,$ then
    for all $f \in \mathcal{C}^{2}_{0}(\mathbb{R}^d \times \mathbb{R})$
    with bounded first and second derivatives
    and for all $\eta \in \mathcal{M}_{f}(\mathbb{R}^d)$, we have

    \begin{equation}\label{eq: gamma - r radius comparison bound}
    \gamma\big(x,\rho_{\gamma}*\eta(x)\big)
    \left| \theta 
        \int_{\mathbb{R}^n} 
            \big\{
                f\big(y,\rho_{r}*\eta(y)\big)-f\big(x,\rho_{r}*\eta(x)\big)
            \big\}
        q_{\theta}(x,dy)
    \right|
    \leq C_{\gamma, f}    
    \end{equation}
    for some constant $C_{\gamma,f}$ independent of $\eta$ and $\theta$. 
\end{lemma}
\begin{proof}
Since $f(x_1,...,x_d,w)$ is twice differentiable, we can Taylor expand $f$ around  $(x_1,...,x_d,\rho_{r}*\eta(x))$ up to second order and obtain that for $x=(x_1,..,x_d)$,
\begin{equation}
\begin{aligned}
&~f\big(y,\rho_{r}*\eta(y)\big)-f\big(x,\rho_{r}*\eta(x)\big)\\
=&~ \nabla_x f \big(x,\rho_{r}*\eta(x)\big)
            \cdot (y-x)\\ 
        &~ + \frac{\partial f}{\partial w}
            \big(x,\rho_{r}*\eta(x)\big) \left(\rho_{r}*\eta(y)-\rho_{r}*\eta(x)\right)\\
        &~ + \frac{1}{2}\sum_{1 \leq i,j \leq d }\frac{\partial^2 f}{\partial x_i x_j}
            \big(z,\rho_{r}*\eta(z)\big) (y_i-x_i)(y_j-x_j)\\
        &~ + \frac{1}{2}\sum_{1 \leq i \leq d}
                \frac{\partial^2 f}{\partial w \partial x_i}
                 \big(z,\rho_{r}*\eta(z)\big) (y_i-x_i)
                    \left(\rho_{r}*\eta(y)-\rho_{r}*\eta(x)\right)\\
        &~ + \frac{1}{2}\frac{\partial^2 f}{\partial^2 w}
            \big(z,\rho_{r}*\eta(z)\big)
            \left(
            \rho_{r}*\eta(y)-\rho_{r}*\eta(x)
            \right)^2,
\end{aligned}    
\end{equation}
where $z=ty+(1-t)x$ for some $t \in [0,1]$.

@@

\begin{align}
J'(s)=&  -\frac{\epsilon^d_r |\covq|}
        {2\left|\covq s +  \epsilon^2_r I \right|^{3/2}}
        \exp \left(
                    \frac{(x-w-\meanq s)^{T}
                        (\covq s +  \epsilon^2_r I)^{-1}
                        (x-w-\meanq s)}{2}
                        -\frac{|x-w|^2}{2\epsilon^2}
            \right)\\
    &+\frac{\epsilon^d_r }
        {\left|\covq s +  \epsilon^2_r I \right|^{1/2}}
        \exp \left(
                    \frac{(x-w-\meanq s)^{T}
                        (\covq s +  \epsilon^2_r I)^{-1}
                        (x-w-\meanq s)}{2}
                        -\frac{|x-w|^2}{2\epsilon^2}
            \right) \\
    & \qquad \times \frac{1}{2}
                    \bigg\{
                         -\meanq^{T}(\covq s + \epsilon^2_r I)^{-1}(x-w-\meanq s)\\
    & \qquad \qquad     -(x-w-\meanq s)^{T}
                        (\covq s +  \epsilon^2_r I)^{-1}\meanq\\
    & \qquad \qquad     -(x-w-\meanq s)^{T}
                        \covq(\covq s +  \epsilon^2_r I)^{-2}(x-w-\meanq s)                    
                    \bigg\},
\end{align}
which is continuous and bounded near $s=0$.
As a result,
\begin{align}
\left| \theta \int_{\mathbb{R}^d}
                \left(p_{\epsilon^2_r}(y-w)
                        -p_{\epsilon^2_r}(x-w)
                \right)
                q_\theta(x,dy)
        \right|
        \leq&  \theta \left(J\left(1 / \theta \right)-1\right) 
                \times \epsilon_\gamma^d \epsilon_r^{-d}
                p_{\epsilon^2_\gamma}(x-w)\\
         =&  J'(u) 
                \times \epsilon_\gamma^d \epsilon_r^{-d}
                p_{\epsilon^2_\gamma}(x-w)        
\end{align}
for some $u \in [0, 1/\theta]$ by Mean Value Theorem. We have therefore established our bound as $J(s)$ has bounded derivative in a neighbourhood around $0$.

@@

\begin{proof}
    \comment{TODO: superscript $\theta$ should be $N$ below}
We first take $f \equiv 1$ in Equation \eqref{eq: Pre-limit Martingale Characterisation}.
\comment{TODO: Write $[M]_t$ instead of $[M_t]$ for quadratic variation.}
We have that 
\begin{equation}
    \label{eq: Total Mass Martingale Characterisation}
\begin{aligned}
     M^{\theta}_t(1):=&  \langle 1,\eta^{N}_t \rangle
        -\langle 1,\eta^{N}_0 \rangle\\
 &-  \int_{0}^{t}\bigg\{   \int\left(   \int \theta
 \left(
        r\big(z,\rho_{r}*\eta^{N}_{s}(z)\big)
            -r\big(x,\rho_{r}*\eta^{N}_{s}(x)\big)
            \right)
                    q_\theta(x,dz)\right)
                            \gamma\big(x,\rho_{\gamma}*\eta^{N}_{s}(x)\big)
                            \eta^{N}_{s}(dx)\\
& \qquad +\int F\big(x,\rho_{F}*\eta^{N}_{s}(x)\big)
\eta^{N}_{s}(dx)\bigg\} ds
\end{aligned}    
\end{equation}
is a martingale.
Therefore, taking expectations, we have
\begin{equation}\label{eq: Total mass control generalised}
\begin{aligned}
&\mathbb{E}[\langle 1 ,\eta^{N}_{t}\rangle] \\
\leq &  \mathbb{E}[\langle 1 ,\eta^{N}_{0}\rangle] \\
&+ \int_{0}^{t}\mathbb{E}\bigg\langle
        \gamma\big(x,\rho_{\gamma}*\eta^{N}_{s}(x)\big)
            \left| \theta \int_{\mathbb{R}^n}
                    \big\{r\big(\rho_{r}*\eta^{N}_s(y)\big)-r\big(\rho_{r}*\eta^{N}_s(x)\big)\big\}
                    \rho_{1/\theta}(x-y)dy\right| + C_F 
                            ,\eta^{N}_s(dx)  \bigg\rangle ds\\
\leq &  \mathbb{E}[\langle 1 ,\eta^{N}_{0}\rangle] + (C_{\gamma,r} + C_F)\int_{0}^{t}\mathbb{E} \langle 1  ,\eta^{N}_s(dx) \rangle ds,
\end{aligned}
\end{equation}
where the last inequality holds by Lemma \ref{lem: gamma - r radius comparison}.

>>>>

\begin{definition}[Model parameterization and assumptions]
\label{def:model_setup}
Suppose $\gamma(x, m)$, $r(x, m)$, and $F(x, m)$
are uniformly continuous functions
from $\IR^d \times \IR_{\ge 0}$ to $\IR$.
Furthermore, suppose $r$ takes values in $[0, 1]$
and has uniformly bounded second derivatives,
that $\gamma(x,m) \ge 0$,
and that $m^2 \gamma(x, m)$ is uniformly bounded.
Fix positive values $\theta$, $\epsilon_r$, $\epsilon_\gamma$, and $\epsilon_F$,
with $\epsilon_r < \epsilon_\gamma$.
Suppose $F(x,m)$ is bounded above
with uniformly bounded first and second derivatives in $m$,
and define
\begin{align} \label{eqn:mu_defn}
    \mu(x, \eta)
    =
    \max\left(0, 
        r(x, \smooth{r} \eta(x))
        \gamma(\smooth{\gamma} \eta(x))
        - \frac{1}{\theta}
        F(x, \smooth{F} \eta(x)) 
    \right) .
\end{align}
Although the models will make sense without this requirement,
for the purposes of convergence we will also require that
$\theta > 2 \sup_x \sup_{y : \|y\| = 1} y^T \covq(x) y/(\epsilon_\gamma^2 - \epsilon_r^2)$.
\end{definition}

@@

\comment{TODO: move actual proof after its ingredients; just foreshadow here?}

\begin{proof}[Proof of Theorem~\ref{thm:nonlocal_convergence}.]
Here, we need to prove that for any sequence $N_k \to \infty$,
there is a subsequence $N_{k(j)}$ along which $(\eta^{N_{k(j)}}_t)_{t\ge 0}$ converges in distribution
(i.e., tightness for the family of distributions on c\`adl\`ag measure-valued paths),
and that any limit is a solution to the martingale problem stated in the Theorem.
This would imply convergence along $N_k$ to a unique limit
if the solution to the martingale problem was unique;
however, we only know this to be the case in certain situations.

@@

To prove tightness of the population process $\eta_t$,
we first show that the expected total mass at some fixed time $T$
is controlled uniformly across $N$, a compact containment condition.
To establish this control,
we first prove a technical lemma which controls spatial density of our population,
and then we use the Burkholder-Davis-Gundy inequality
to give us control over the running supremum of the total population mass.
Once we have control over the total mass of the population, 
we will also have control over local densities.
This then allows us to prove that the particle system is tight
by an application of the Aldous-Rebolledo criterion \citet{rebolledo:1980,alisonsup}.

% % % % % % % % % % % % %
\subsubsection{Compact Containment Condition}

To show compact containment, we will need boundedness
of the total net reproduction rate.
This does not follow immediately from our assumption that $F$ is bounded above
because reproduction is not local.
\comment{say more about the difference between $F$ and the total net reproduction rate}
The following lemma will be useful.

\begin{lemma}
    \label{lem:q_bound}
    Under the assumptions of Definition~\ref{def:model_setup},
    for all $f \in \mathcal{C}^{2}_{0}(\IR^d \times [0,\infty))$,
    then there exists a $C_f < \infty$ such that
    for all $\theta > 0$ and all $\eta \in \measures$,
    \begin{equation*}
        \left| \theta 
            \int_{\IR^n} 
                \left\{
                    f\left(y, \smooth{r}\eta(y)\right)
                    -
                    f\left(x, \smooth{r}\eta(x)\right)
                \right\}
            q_{\theta}(x,dy)
        \right|
        \leq
        C_f \left( 1 + \smooth{\gamma} \eta(x)^2 \right)
    \end{equation*}
\end{lemma}

\begin{proof}[of Lemma \ref{lem:q_bound}]
    The key assumption here from Definition~\ref{def:model_setup}
    will be that $\epsilon_{r}^2 + \frac{2\lambda}{\theta} < \epsilon_{\gamma}^2$,
    where $\lambda = \sup_x \sup_{y : \|y\| = 1} y^T \covq(x) y$
    is the largest eigenvalue of $\covq(x)$ across all $x$.
    First note that 
    \begin{align} \label{eqn:q_bound_split}
    &
        \left| \theta 
            \int_{\IR^n} 
                \left\{
                    f\left(y, \smooth{r}\eta(y)\right)
                    -
                    f\left(x, \smooth{r}\eta(x)\right)
                \right\}
            q_{\theta}(x,dy)
        \right|
    \\ &\qquad {} \leq
        \left| \theta 
            \int_{\IR^n} 
                \left\{
                    f\left(y, \smooth{r}\eta(y)\right)
                    -
                    f\left(x, \smooth{r}\eta(y)\right)
                \right\}
            q_{\theta}(x,dy)
        \right|
    \\ &\qquad \qquad {}  +
        \left| \theta 
            \int_{\IR^n} 
                \left\{
                    f\left(x, \smooth{r}\eta(y)\right)
                    -
                    f\left(x, \smooth{r}\eta(x)\right)
                \right\}
            q_{\theta}(x,dy) 
        \right| .
    \end{align}

    Since 
    \begin{align*}
        f\left(y, \smooth{r}\eta(y)\right) - f\left(x, \smooth{r}\eta(y)\right)
        &=
        \sum_i (y_i - x_i) \partial_{x_i} f \left(x, \smooth{r}\eta(y)\right) 
        \\ &\qquad {}
        +
        \frac{1}{2} \sum_{ij} (y_i - x_i)(y_j - x_j)
            \partial_{x_i x_j} f \left(z, \smooth{r}\eta(y)\right) 
    \end{align*}
    where $z = t x + (1-t) y$ for some $0 \le t \le 1$,
    writing
    $K_1 = \sup_{x,m} \max_i |\partial_{x_i} f(x, m)|$
    and 
    $K_2 = \sup_{x,m} \max_{i,j} |\partial_{x_i x_j} f(x, m)|$
    for the bounds on the spatial derivatives of $f$,
    the first term is bounded by
    \begin{align*}
        K_1 \left|
            \theta \int_{\IR^d} \sum_i (y_i - x_i) q_\theta(x, dy)
        \right|
        +
        K_2 \left|
            \theta \int_{\IR^d} \sum_{ij} (y_i - x_i) (y_j - x_j)  q_\theta(x, dy)
        \right| .
    \end{align*}
    This is bounded uniformly because
    $q_\theta(x, dy)$ is the density of a Gaussian with mean $\meanq(x)/\theta$
    and covariance $\covq(x)/\theta$, and both $\meanq(x)$ and $\covq(x)$
    are bounded uniformly in $x$.

    Similarly, 
    \begin{align*}
        f(x, \smooth{r}\eta(y)) - f(x, \smooth{r}\eta(x))
        &=
        (\smooth{r}\eta(y) - \smooth{r}\eta(x)) f'(x, \smooth{r}\eta(x))
        + \frac{1}{2} (\smooth{r}\eta(y) - \smooth{r}\eta(x))^2 f''(x, m)
    \end{align*}
    for some $\smooth{r}\eta(x) \le m \le \smooth{r}\eta(y)$,
    where $f'$ and $f''$ denote the first and second derivatives with respect to the second argument.
    So,
    the second term in~\eqref{eqn:q_bound_split} is bounded by
    \begin{align*}
        \|f'\|_\infty
        \left|
            \theta \int_{\IR^d}
                (\smooth{r}\eta(y) - \smooth{r}\eta(x))
            q_\theta(x, dy)
        \right|
        +
        \|f''\|_\infty
        \left|
            \theta \int_{\IR^d}
                (\smooth{r}\eta(y) - \smooth{r}\eta(x))^2
            q_\theta(x, dy)
        \right| .
    \end{align*}
    Therefore, it suffices to show that there is a $K < \infty$ such that
    for all $\eta \in \measures$ and $x \in \IR^d$,
    \begin{equation}
        \label{eqn:first_moment_rho}
        \left|\theta \int_{\mathbb{R}^d}
            ( \smooth{r} \eta(y)-\smooth{r} \eta(x) )
        q_\theta(x,dy)\right|
        \leq K \smooth{\gamma} \eta(x)
    \end{equation}
    and
    \begin{equation}
        \label{eqn:second_moment_rho}
        \theta \int_{\mathbb{R}^d}
                \left( \smooth{r} \eta(y) - \smooth{r} \eta(x) \right)^2
        q_\theta(x,dy)
        \leq K \left(\smooth{\gamma}\eta(x) \right)^2 .
    \end{equation}
    Note that the right hand side of each is the average density
    over a wider region (since $\epsilon_\gamma > \epsilon_r$).

    First we prove \eqref{eqn:first_moment_rho}.
    Recall that $\smooth{r} \eta(x) = \int p_{\epsilon^2_r}(x - w) \eta(dw)$,
    where $p_t$ is the density of a Gaussian with mean 0 and variance $t$,
    so that applying Fubini, \eqref{eqn:first_moment_rho} is
    \begin{align*}
        \left|
        \int_{\IR^d}
        \theta
            \int_{\IR^d}
                ( p_{\epsilon_r^2}(y - w) - p_{\epsilon_r^2}(x - w) )
            q_\theta(x, dy)
        \eta(dw)
        \right| .
    \end{align*}
    Write $p_{s, x}(\cdot)$ for the density of a Gaussian
    with mean $s \meanq(x)$ and covariance $\epsilon_r^2 I + s \covq(x)$,
    so that $\int p_{\epsilon_r^2}(y - w) q_\theta(x, dy) = p_{\epsilon_r^2, 1/\theta, x}(w-x)$.
    It therefore suffices to show that for all $x$ and $w \in \IR^d$,
    \begin{align*}
        \left|
        \theta
            \int_{\IR^d}
                ( p_{\epsilon_r^2}(y-w) - p_{\epsilon_r^2}(x-w) )
            q_\theta(x, dy)
        \right|
        &=
        \theta \left|
            p_{1/\theta, x}(w-x) - p_{0, x}(w-x)
        \right|
        \le
        p_{\epsilon^2_\gamma}(w-x) .
    \end{align*}
    However,
    $\theta( p_{1/\theta, x}(z) - p_{0, x}(z) )
    = \partial_s p_{s, x}(z)$ for some $0 \le s \le 1/\theta$.
    Write $\Gamma(s,x) = \epsilon_r^2 I + s \covq(x)$,
    so that
    \begin{align*}
        p_{s, x}(z)
        &=
        \frac{1}{\left(2 \pi |\Gamma(s,x)|\right)^{d/2}}
        \exp\left(
            -\frac{1}{2} (z - s\meanq(x))^T \Gamma(s,x)^{-1} (z - s\meanq(x))
        \right) ,
    \end{align*}
    and note that if $\lambda_i$ are the eigenvalues of $\covq(x)$
    then $|\Gamma(s, x)| = \prod_i (\epsilon_r^2 + s \lambda_i)$,
    and
    $\partial_s |\Gamma(s, x)| = \sum_i \lambda_i |\Gamma(s, x)| / (\epsilon_r^2 + s \lambda_i)$.
    Therefore,
    \begin{align*}
        \partial_s p_{s, x}(z)
        &=
        \left(
            \meanq(x)^T \Gamma(s,x)^{-1} (z - s \meanq)
            +
            (z - s \meanq)^T
            \Gamma(s,x)^{-1} \covq(x) \Gamma(s,x)^{-1}
            (z - s \meanq)
            -
            \sum_i \frac{\lambda_i}{\epsilon_r^2 + s \lambda_i}
        \right)
        p_{s, x}(z) ,
    \end{align*}
    where $z^T$ is the transpose of $z$.
    Since $\partial_s p_{s, x}(z) / p_{s, x}(z)$ is a polynomial in $z$
    with coefficients that are uniformly bounded for $s \ge 1$,
    it is evident that
    $\sup_z \partial_s p_{\epsilon_r^2, s, x}(z) / p_{\epsilon_\gamma^2}(z) < \infty$
    if and only if
    \begin{align*}
        \frac{1}{2\epsilon_\gamma^2} \|z\|^2
        -\frac{1}{2} (z - s\meanq(x))^T \Gamma(s,x)^{-1} (z - s\meanq(x))
    \end{align*}
    is negative for all $z$ outside a bounded region.
    Since $\inf_z z^T \Gamma(s,x)^{-1} z / \|z\|^2 = 1 / (s \lambda(x) + \epsilon_r^2)$,
    where $\lambda(x) = \sup z^T \covq(x) z / \|z\|^2$ is the largest eigenvalue of $\covq(x)$,
    equation~\eqref{eqn:first_moment_rho} follows from the assumption
    that $\epsilon_r^2 + 2 \sup_x \lambda(x)/\theta < \epsilon_\gamma^2$.
    (Note that we do not yet need the factor of 2.)

    Next we prove equation~\eqref{eqn:second_moment_rho}, in a similar way.
    Again applying Fubini,
    \begin{align*}
        &
        \theta \int_{\mathbb{R}^d}
                \left( \smooth{r} \eta(y) - \smooth{r} \eta(x) \right)^2
        q_\theta(x,dy)
        \\ &\qquad
        =
        \int_{\IR^d} \int_{\IR^d}
        \theta \int_{\IR^d}
            ( p_{\epsilon_r^2}(y - w) - p_{\epsilon_r^2}(x - w) )
            ( p_{\epsilon_r^2}(y - v) - p_{\epsilon_r^2}(x - v) )
        q_\theta(x, dy)
        \eta(dv) \eta(dw) ,
    \end{align*}
    and so as before, equation~\eqref{eqn:second_moment_rho} will follow if
    the integrand is bounded by $K p_\gamma(x-w) p_\gamma(x - v)$.
    % \begin{align*}
    %     \theta \int_{\IR^d}
    %         ( p_{\epsilon_r^2}(y - w) - p_{\epsilon_r^2}(x - w) )
    %         ( p_{\epsilon_r^2}(y - v) - p_{\epsilon_r^2}(x - v) )
    %     q_\theta(x, dy)
    %     \le K p_\gamma(x-w) p_\gamma(x - v) .
    % \end{align*}
    Now, let $Y_1$, $Y_2$, and $Z$ be independent $d$-dimensional Gaussians with mean zero,
    where $Y_1$ and $Y_2$ have covariance $\sigma^2_r I$,
    and $Z$ has covariance $\covq(x)$.
    Write $p_{s,t,x}(\cdot, \cdot)$ for the joint density of
    $Y_1 + \sqrt{s} Z + s \meanq(x)$ and $Y_2 + \sqrt{t} Z + t \meanq(x)$.
    Then, observe that
    \begin{align*}
        &
        \theta \int_{\IR^d}
            ( p_{\epsilon_r^2}(y - w) - p_{\epsilon_r^2}(x - w) )
            ( p_{\epsilon_r^2}(y - v) - p_{\epsilon_r^2}(x - v) )
        q_\theta(x, dy)
        \\ &\qquad
        =
        \theta \left(
            p_{1/\theta, 1/\theta, x}(x-w, x-v)
            - p_{0, 1/\theta, x}(x-w, x-v)
        \right. \\ &\qquad \qquad \left. {}
            - p_{1/\theta, 0, x}(x-w, x-v)
            + p_{0, 0, x}(x-w, x-v)
        \right) 
        \\ &\qquad 
        =
            \partial_s p_{s, 1/\theta, x}(x-w, x-v)
            - \partial_t p_{0, t, x}(x-w, x-v) ,
    \end{align*}
    for some $0 \le s, t \le 1/\theta$.
    As before, $\partial_s p_{s, t, x}(y_1, y_2)$ has polynomial terms multiplying
    $$
        \exp\left(
            - \frac{1}{2} [y_1, y_2[^T \Gamma(s, t, x)^{-1} [y_1, y_2]
        \right) ,
    $$
    where $[y_1, y_2]$ is the $\IR^{2d}$ vector formed by concatenating $y_1$ and $y_2$,
    and $\Gamma(s, t, x)$ is the block matrix
    \begin{align*}
        \Gamma(s, t, x)
        =
        \left[
        \begin{array}{cc}
            \epsilon_r^2 I + s \covq(x) & \sqrt{st} \covq(x) \\
            \sqrt{st} \covq(x) & \epsilon_r^2 I + t \covq(x) \\
        \end{array}
        \right] .
    \end{align*}
    If $\covq(x) u = a u$ for some $a \in \IR$,
    then $[u \sqrt{s}, u \sqrt{t}]$ is an eigenvector of $\Gamma(s, t, x)$
    with eigenvalue $\epsilon^2_r + (s+t) a$,
    and $[u \sqrt{t}, - u \sqrt{s}]$ is an eigenvector of $\Gamma(s, t, x)$
    with eigenvalue 0.
    This implies the largest eigenvalue of $\Gamma(s, t, x)$
    is equal to $\epsilon^2_r + (s+t) \lambda(x)$,
    where $\lambda(x)$ is again the largest eigenvalue of $\covq(x)$.
    Therefore, if $s + t \le 2 / \theta$,
    \begin{align*}
        &
        (\|y_1\|^2 + \|y_2\|^2) / \epsilon^2_\gamma
            - [y_1, y_2[^T \Gamma(s, t, x)^{-1} [y_1, y_2]
        \\ &\qquad \le
        (\|y_1\|^2 + \|y_2\|^2) \left(
            \frac{1}{\epsilon^2_\gamma}
            - \frac{1}{\epsilon^2_r +  2\lambda(x)/\theta}
        \right) ,
    \end{align*}
    which is negative by assumption.
    Therefore, there is a $K$ such that
    \begin{align*}
        \frac{ \left|
            \partial_s p_{s, 1/\theta, x}(x-w, x-v)
            - \partial_t p_{0, t, x}(x-w, x-v)
        \right| }{
            p_{\epsilon^2_\gamma}(x - w)
            p_{\epsilon^2_\gamma}(x - v)
        }
        \le K
    \end{align*}
    for all $\theta > 1$ and all $x$, $v$, and $w \in \IR^d$,
    proving equation~\eqref{eqn:second_moment_rho} and hence the lemma.
\end{proof}

\begin{corollary}
    \label{cor:gamma_bound}
    Under the assumptions of Definition~\ref{def:model_setup},
    including that $\epsilon_{r}^2 + \frac{1}{\theta} < \epsilon_{\gamma}^2$,
    for all $f \in \mathcal{C}^{2}_{0}(\IR^d \times \IR)$
    with bounded first and second derivatives
    and for all $\eta \in \measures$, we have
    \begin{equation}\label{eq: gamma - r radius comparison bound}
    \gamma\big(x,\rho_{\gamma}*\eta(x)\big)
    \left| \theta 
        \int_{\IR^n} 
            \big\{
                f\big(y,\rho_{r}*\eta(y)\big)-f\big(x,\rho_{r}*\eta(x)\big)
            \big\}
        q_{\theta}(x,dy)
    \right|
    \leq C_{\gamma, f}    
    \end{equation}
    for some constant $C_{\gamma,f}$ independent of $\eta$ and $\theta$. 
\end{corollary}

\begin{proof}
Since $f(x_1,...,x_d,w)$ is twice differentiable, we can Taylor expand $f$ around  $(x_1,...,x_d,\rho_{r}*\eta(x))$ up to second order and obtain that for $x=(x_1,..,x_d)$,
\begin{equation}
\begin{aligned}
&~f\big(y,\rho_{r}*\eta(y)\big)-f\big(x,\rho_{r}*\eta(x)\big)\\
=&~ \nabla_x f \big(x,\rho_{r}*\eta(x)\big)
            \cdot (y-x)\\ 
        &~ + \frac{\partial f}{\partial w}
            \big(x,\rho_{r}*\eta(x)\big) \left(\rho_{r}*\eta(y)-\rho_{r}*\eta(x)\right)\\
        &~ + \frac{1}{2}\sum_{1 \leq i,j \leq d }\frac{\partial^2 f}{\partial x_i x_j}
            \big(z,\rho_{r}*\eta(z)\big) (y_i-x_i)(y_j-x_j)\\
        &~ + \frac{1}{2}\sum_{1 \leq i \leq d}
                \frac{\partial^2 f}{\partial w \partial x_i}
                 \big(z,\rho_{r}*\eta(z)\big) (y_i-x_i)
                    \left(\rho_{r}*\eta(y)-\rho_{r}*\eta(x)\right)\\
        &~ + \frac{1}{2}\frac{\partial^2 f}{\partial^2 w}
            \big(z,\rho_{r}*\eta(z)\big)
            \left(
            \rho_{r}*\eta(y)-\rho_{r}*\eta(x)
            \right)^2,
\end{aligned}    
\end{equation}
where $z=ty+(1-t)x$ for some $t \in [0,1]$.

@@

\begin{align}
J'(s)=&  -\frac{\epsilon^d_r |\covq|}
        {2\left|\covq s +  \epsilon^2_r I \right|^{3/2}}
        \exp \left(
                    \frac{(x-w-\meanq s)^{T}
                        (\covq s +  \epsilon^2_r I)^{-1}
                        (x-w-\meanq s)}{2}
                        -\frac{|x-w|^2}{2\epsilon^2}
            \right)\\
    &+\frac{\epsilon^d_r }
        {\left|\covq s +  \epsilon^2_r I \right|^{1/2}}
        \exp \left(
                    \frac{(x-w-\meanq s)^{T}
                        (\covq s +  \epsilon^2_r I)^{-1}
                        (x-w-\meanq s)}{2}
                        -\frac{|x-w|^2}{2\epsilon^2}
            \right) \\
    & \qquad \times \frac{1}{2}
                    \bigg\{
                         -\meanq^{T}(\covq s + \epsilon^2_r I)^{-1}(x-w-\meanq s)\\
    & \qquad \qquad     -(x-w-\meanq s)^{T}
                        (\covq s +  \epsilon^2_r I)^{-1}\meanq\\
    & \qquad \qquad     -(x-w-\meanq s)^{T}
                        \covq(\covq s +  \epsilon^2_r I)^{-2}(x-w-\meanq s)                    
                    \bigg\},
\end{align}
which is continuous and bounded near $s=0$.
As a result,
\begin{align}
\left| \theta \int_{\mathbb{R}^d}
                \left(p_{\epsilon^2_r}(y-w)
                        -p_{\epsilon^2_r}(x-w)
                \right)
                q_\theta(x,dy)
        \right|
        \leq&  \theta \left(J\left(1 / \theta \right)-1\right) 
                \times \epsilon_\gamma^d \epsilon_r^{-d}
                p_{\epsilon^2_\gamma}(x-w)\\
         =&  J'(u) 
                \times \epsilon_\gamma^d \epsilon_r^{-d}
                p_{\epsilon^2_\gamma}(x-w)        
\end{align}
for some $u \in [0, 1/\theta]$ by Mean Value Theorem. We have therefore established our bound as $J(s)$ has bounded derivative in a neighbourhood around $0$.
\comment{This is not enough, since we need to show the constant holds for all $x$, and $J$ depends on $x$.}

@@

\begin{proof}
    \comment{TODO: superscript $\theta$ should be $N$ below}
We first take $f \equiv 1$ in Equation \eqref{eq: Pre-limit Martingale Characterisation}.
\comment{TODO: Write $[M]_t$ instead of $[M_t]$ for quadratic variation.}
We have that 
\begin{equation}
    \label{eq: Total Mass Martingale Characterisation}
\begin{aligned}
     M^{\theta}_t(1):=&  \langle 1,\eta^{N}_t \rangle
        -\langle 1,\eta^{N}_0 \rangle\\
 &-  \int_{0}^{t}\bigg\{   \int\left(   \int \theta
 \left(
        r\big(z,\rho_{r}*\eta^{N}_{s}(z)\big)
            -r\big(x,\rho_{r}*\eta^{N}_{s}(x)\big)
            \right)
                    q_\theta(x,dz)\right)
                            \gamma\big(x,\rho_{\gamma}*\eta^{N}_{s}(x)\big)
                            \eta^{N}_{s}(dx)\\
& \qquad +\int F\big(x,\rho_{F}*\eta^{N}_{s}(x)\big)
\eta^{N}_{s}(dx)\bigg\} ds
\end{aligned}    
\end{equation}
is a martingale.
Therefore, taking expectations, we have
\begin{equation}\label{eq: Total mass control generalised}
\begin{aligned}
&\mathbb{E}[\langle 1 ,\eta^{N}_{t}\rangle] \\
\leq &  \mathbb{E}[\langle 1 ,\eta^{N}_{0}\rangle] \\
&+ \int_{0}^{t}\mathbb{E}\bigg\langle
        \gamma\big(x,\rho_{\gamma}*\eta^{N}_{s}(x)\big)
            \left| \theta \int_{\mathbb{R}^n}
                    \big\{r\big(\rho_{r}*\eta^{N}_s(y)\big)-r\big(\rho_{r}*\eta^{N}_s(x)\big)\big\}
                    \rho_{1/\theta}(x-y)dy\right| + C_F 
                            ,\eta^{N}_s(dx)  \bigg\rangle ds\\
\leq &  \mathbb{E}[\langle 1 ,\eta^{N}_{0}\rangle] + (C_{\gamma,r} + C_F)\int_{0}^{t}\mathbb{E} \langle 1  ,\eta^{N}_s(dx) \rangle ds,
\end{aligned}
\end{equation}
where the last inequality holds by Lemma \ref{cor:gamma_bound}.
